---
title: "Lab 3"
author: "Xuan Wen Loo"
date: "2023-03-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Link to github repo: https://github.com/xuanwenloo/Homework7 

```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)
```

```{r}
data(happy, package="productplots")
head(happy)
```

```{r}
HAPPY <- readRDS("data/HAPPY.rds")
```

**Part one (cleaning)**

1. Data cleaning: the values “IAP”, “DK” and “NA” all encode missing values. Replace all of these instances by the value NA.
```{r}
HAPPY[HAPPY == "IAP" | HAPPY == "DK"| HAPPY == 'NA'] = NA 
head(HAPPY)
```

2. Check the type of the variable and cast into the right type (factor variable for categorical variables). For age, change “89 OR OLDER” to 89 and assume the variable should be numeric.
```{r}
str(HAPPY)
```

```{r}
HAPPY$AGE[HAPPY$AGE == '89 OR OLDER'] = 89
HAPPY$AGE = as.numeric(HAPPY$AGE)
str(HAPPY)
```

3. Bring all levels of factors into a sensible order. For marital you could e.g. order the levels according to average age.
```{r}
col = c('HAPPY', 'SEX', 'MARITAL', 'DEGREE', 'FINRELA', 'HEALTH', 'PARTYID', 'POLVIEWS')
HAPPY[col] = lapply(HAPPY[col], factor)
levels(HAPPY$MARITAL)
levels(HAPPY$DEGREE)
levels(HAPPY$FINRELA)
levels(HAPPY$HEALTH)
levels(HAPPY$PARTYID)
levels(HAPPY$POLVIEWS)
```

```{r}
h = HAPPY
h$MARITAL = factor(h$MARITAL, levels = c('NEVER MARRIED', 'MARRIED', 'SEPARATED', 'DIVORCED', 'WIDOWED'))

h$DEGREE = factor(h$DEGREE, levels = c('LT HIGH SCHOOL', 'HIGH SCHOOL', 'JUNIOR COLLEGE', 'BACHELOR', 'GRADUATE'))

h$FINRELA = factor(h$FINRELA, levels = c('FAR BELOW AVERAGE', 'BELOW AVERAGE', 'AVERAGE', 'ABOVE AVERAGE', 'FAR ABOVE AVERAGE'))

h$HEALTH = factor(h$HEALTH, levels = c('POOR', 'FAIR', 'GOOD', 'EXCELLENT'))

h$PARTYID = factor(h$PARTYID, levels = c('STRONG DEMOCRAT', 'NOT STR DEMOCRAT', 'IND,NEAR DEMOCRAT', 'INDEPENDENT', 'IND,NEAR REP', 'NOT STR REPUBLICAN', 'STRONG REPUBLICAN', 'OTHER PARTY'))

h$POLVIEWS = factor(h$POLVIEWS, levels = c('EXTREMELY LIBERAL', 'LIBERAL', 'SLIGHTLY LIBERAL', 'MODERATE', 'SLIGHTLY CONSERVATIVE', 'CONSERVATIVE', 'EXTREMLY CONSERVATIVE'))

HAPPY = h
str(HAPPY)
```

```{r}
ggplot(HAPPY, aes(x = HAPPY)) + geom_bar()
```
```{r}
ggplot(HAPPY, aes(x = YEAR)) + geom_bar()
```
```{r}
ggplot(HAPPY, aes(x = AGE)) + geom_bar()
```
```{r}
ggplot(HAPPY, aes(x = SEX)) + geom_bar()
```
```{r}
ggplot(HAPPY, aes(x = MARITAL)) + geom_bar()
```
```{r}
ggplot(HAPPY, aes(x = DEGREE)) + geom_bar()
```
```{r}
ggplot(HAPPY, aes(x = FINRELA)) + geom_bar()
```
```{r}
ggplot(HAPPY, aes(x = HEALTH)) + geom_bar()
```
```{r}
ggplot(HAPPY, aes(x = PARTYID)) + geom_bar()
```
```{r}
ggplot(HAPPY, aes(x = POLVIEWS)) + geom_bar()
```

**Part two (individual exploration)**

I will be investigating the relationship between happiness and degree and marital status.

```{r}
ggplot(na.omit(HAPPY), aes(x = DEGREE, fill = HAPPY)) + geom_bar(position = "fill")
```

The plot shows that people with higher academic qualifications are happier, most probably because they have higher achievements in the society, and they can get a better job with a good pay.

```{r}
ggplot(na.omit(HAPPY), aes(x = MARITAL, fill = HAPPY)) + geom_bar(position = "fill")
```

The plot shows that the proportion of very happy people is the largest in married couples, and least among separated couples. This is most probably because married people have a harmonious family which will make them happier.

```{r, message=FALSE, warning=FALSE}
library(ggmosaic)
ggplot(na.omit(HAPPY)) + geom_mosaic(aes(x = product(DEGREE), fill = HAPPY)) + facet_wrap(~MARITAL) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

From the plot, it seemed that married people with high academic achievements are most likely to be happier compared to people who failed a relationship and did not have high degrees.
